`State and Parameter Estimation for Spatio-Temporal Neural Fields' or `A Data Driven Framework for Patient-Specific Neural Field Modelling'

### 2010-06-17 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
corrected fig8 caption
random typo correction

Histograms of the parameter estimates obtained over the 100 realisations are shown in Figure~\ref{fig:Parameters}. The actual parameter values, the mean estimates, corresponding standard deviations and biases are given in Table~\ref{tab:Parameters estimates}. The biases show the percentage differences between the actual values and the mean parameter estimates. The mean estimates are in good concordance with the actual values. 
% It should be noted that when the excitatory connectivity parameters, $\hat\theta_0$ and $\hat\theta_2$, were under (or over) estimated, the inhibitory connectivity parameter, $\hat\theta_1$, tended to be over (or under) estimated. This is demonstrated in Figure~\ref{fig:ParametersRatio} where we show the distribution of the ratios of the parameter estimates. 
\begin{table}
\centering
\begin{tabular}{cccc}
	\hline\hline Parameter & Actual Value & Mean Estimate $\pm$ Std. Dev. & Bias (\%) \\
	\hline\hline
 	$\xi$ & $0.9$ & $0.925\pm 0.004$ & $2.778$ \\
	$\theta_0$ & $10$ & $10.228\pm2.794$ & $2.28$\\
	$\theta_1$ & $-8$ & $-8.495\pm1.913$ & $-6.187$\\
	$\theta_2$ & $0.5$ & $0.545\pm0.077$ & $9$\\
	$\mid\theta_0/\theta_1\mid$ &$1.25 $ & $1.189\pm 0.0812$ &$4.88$\\
	$\theta_0/\theta_2$ &$20$ & $18.567\pm3.757$ &$7.165$ \\
\hline\hline
\end{tabular}
\caption{Actual parameter values compared to parameter estimates.}
\label{tab:Parameters estimates}
\end{table}
\begin{figure}
    \centering
    \subfigure[][]{\includegraphics{./Graph/zeta.pdf}}
    \subfigure[][]{\includegraphics{./Graph/theta0.pdf}}\\
    \subfigure[][]{\label{fig:theta1dis}\includegraphics{./Graph/theta1.pdf}}
    \subfigure[][]{\includegraphics{./Graph/theta2.pdf}}
\caption{Distribution of the parameter estimates over 100
realisations. The dotted line shows the actual value of the parameters and the solid line shows the mean of the estimated parameters. \textbf{a.} The distribution of the $\xi$ estimates where $\xi=1-\zeta T_s $ and $\zeta$ is the reciprocal of the synaptic time constant. \textbf{b.} The distribution of the central excitatory connectivity kernel basis function parameter estimates, $\theta_0$. \textbf{c.} The distribution of the surround inhibition connectivity kernel basis function parameter estimates, $\theta_1$. \textbf{d.} The distribution of the longer range excitatory connectivity kernel basis function parameter estimates, $\theta_2$.}
\label{fig:Parameters}
\end{figure}
\begin{figure}
    \centering
\subfigure[][]{\includegraphics{./Graph/theta0_theta1_ratio.pdf}}
\subfigure[][]{\includegraphics{./Graph/theta0_theta2_ratio.pdf}}\\
\caption{Distribution of the connectivity kernel basis function parameter ratios over 100 realisations. \textbf{a.} The absolute value of the ratio of the local excitatory parameter to the surround inhibition parameter (true ratio is 1.25). \textbf{b.} The ratio of the local excitatory parameter to the longer range excitatory parameter (true ratio is 20). In each case, the solid and dotted lines show the locations of the true and estimated ratios respectively.}
\label{fig:ParametersRatio}
\end{figure}
The reconstructed connectivity kernels from the estimated parameters are illustrated in Figure~\ref{fig:KernelEstimates}, where the mean over the 100 realisations is shown as a solid line and the range indicated by the shaded region, which is  the $95\%$ confidence interval. The figure verifies that the shape of the estimated connectivity kernel is in good accordance to the true kernel. The true kernel lies within the confidence interval; however, the mean estimated kernel is more negative than the actual kernel due to the small bias on the estimates of the inhibitory parameter, $\hat\theta_1$, as seen in Figure~\ref{fig:theta1dis}.

\begin{figure}
    \centering
% \subfigure[][]{
\includegraphics{./Graph/KernelEstimate.pdf}%}
% \subfigure[][]{\includegraphics{./Graph/KernelEstimateFFT.pdf}}\\
\caption{Connectivity kernel estimate. The actual kernel is shown by the solid line, the mean kernel estimate (over 100 realisations) is shown by the dotted line and the $95~\%$ confidence interval is shown by the shaded grey region.}
\label{fig:KernelEstimates}
\end{figure}
The accuracy of the state estimates was evaluated by comparing the field reconstruction to the true field using the mean (over time) of the root mean square error (MRMSE) (over space). The RMSE of the field estimates over time for 100 realisations is shown in Figure~\ref{fig:RMSE}, with the $95~\%$ confidence interval. The error is consistently around 0.5 over time, which approximately $7.1$~\% of the voltages, which range over $(-3.29,3.77)$. The error in the reconstructed field is mostly due to the loss of high spatial frequency information due to the basis function decomposition; this is illustrated in Figure~\ref{fig:FieldEstimate}, which shows a snapshot at a single time instant of the true neural field, the estimated field and the corresponding error, which is the difference between the true and estimated neural fields. It can be seen in Figure~\ref{fig:FieldEstimate}(b) that the estimated field contains the basic structure of the true field (Figure~\ref{fig:FieldEstimate}(a)), but is missing the higher-frequency detail.
  \begin{figure}
   	\begin{center}
   		\includegraphics{./Graph/MRMSE.pdf} 
   	\end{center}
   	\caption{The mean of average RMSE of the estimated field over 100 realisations (solid line) and the 95~\% confidence interval (grey region).} 
\label{fig:RMSE}
   \end{figure}

@@

The closest competitor to the integro-difference equation based approach is perhaps the multivariate auto-regressive (MVAR) model. An advantage of the MVAR-based methods is that they allow for anisotropic connectivity, which is present in cortico-cortical fibres. A recent theoretical study~\cite{Jirsa2009} has demonstrated the importance of this multi-scale structure in generating the characteristic rhythms of the brain. A possible method of creating more precise models with this multi-scale structure maybe to estimate anisotropic cortico-cortical connectivity using a MVAR approach and intracortical connectivity using the methods proposed in this paper. In addition to estimating connectivity independently of the number of observations, the advantage of the methodology presented in this paper, is that the number of parameters required to model the system is considerably less, and the parameter have physical meaning.

>>>>

Histograms of the parameter estimates obtained over the 100 realisations are shown in Figure~\ref{fig:Parameters}. The actual parameter values, the mean estimates, corresponding standard deviations and biases are given in Table~\ref{tab:Parameters estimates}. The biases show the percentage differences between the actual values and the mean parameter estimates. The mean estimates are in good concordance with the actual values. 
% It should be noted that when the excitatory connectivity parameters, $\hat\theta_0$ and $\hat\theta_2$, were under (or over) estimated, the inhibitory connectivity parameter, $\hat\theta_1$, tended to be over (or under) estimated. This is demonstrated in Figure~\ref{fig:ParametersRatio} where we show the distribution of the ratios of the parameter estimates. 
\begin{table}
\centering
\begin{tabular}{cccc}
	\hline\hline Parameter & Actual Value & Mean Estimate $\pm$ Std. Dev. & Bias (\%) \\
	\hline\hline
 	$\xi$ & $0.9$ & $0.925\pm 0.004$ & $2.778$ \\
	$\theta_0$ & $10$ & $10.228\pm2.794$ & $2.28$\\
	$\theta_1$ & $-8$ & $-8.495\pm1.913$ & $-6.187$\\
	$\theta_2$ & $0.5$ & $0.545\pm0.077$ & $9$\\
	$\mid\theta_0/\theta_1\mid$ &$1.25 $ & $1.189\pm 0.0812$ &$4.88$\\
	$\theta_0/\theta_2$ &$20$ & $18.567\pm3.757$ &$7.165$ \\
\hline\hline
\end{tabular}
\caption{Actual parameter values compared to parameter estimates.}
\label{tab:Parameters estimates}
\end{table}
\begin{figure}
    \centering
    \subfigure[][]{\includegraphics{./Graph/zeta.pdf}}
    \subfigure[][]{\includegraphics{./Graph/theta0.pdf}}\\
    \subfigure[][]{\label{fig:theta1dis}\includegraphics{./Graph/theta1.pdf}}
    \subfigure[][]{\includegraphics{./Graph/theta2.pdf}}
\caption{Distribution of the parameter estimates over 100
realisations. The dotted line shows the actual value of the parameters and the solid line shows the mean of the estimated parameters. \textbf{a.} The distribution of the $\xi$ estimates where $\xi=1-\zeta T_s $ and $\zeta$ is the reciprocal of the synaptic time constant. \textbf{b.} The distribution of the central excitatory connectivity kernel basis function parameter estimates, $\theta_0$. \textbf{c.} The distribution of the surround inhibition connectivity kernel basis function parameter estimates, $\theta_1$. \textbf{d.} The distribution of the longer range excitatory connectivity kernel basis function parameter estimates, $\theta_2$.}
\label{fig:Parameters}
\end{figure}
\begin{figure}
    \centering
\subfigure[][]{\includegraphics{./Graph/theta0_theta1_ratio.pdf}}
\subfigure[][]{\includegraphics{./Graph/theta0_theta2_ratio.pdf}}\\
\caption{Distribution of the connectivity kernel basis function parameter ratios over 100 realisations. \textbf{a.} The absolute value of the ratio of the local excitatory parameter to the surround inhibition parameter (true ratio is 1.25). \textbf{b.} The ratio of the local excitatory parameter to the longer range excitatory parameter (true ratio is 20). In each case, the solid and dotted lines show the locations of the true and estimated ratios respectively.}
\label{fig:ParametersRatio}
\end{figure}
The reconstructed connectivity kernels from the estimated parameters are illustrated in Figure~\ref{fig:KernelEstimates}, where the mean over the 100 realisations is shown as a dotted line and the range indicated by the shaded region, which is  the $95\%$ confidence interval. The figure verifies that the shape of the estimated connectivity kernel is in good accordance to the true kernel. The true kernel lies within the confidence interval; however, the mean estimated kernel is more negative than the actual kernel due to the small bias on the estimates of the inhibitory parameter, $\hat\theta_1$, as seen in Figure~\ref{fig:theta1dis}.

\begin{figure}
    \centering
% \subfigure[][]{
\includegraphics{./Graph/KernelEstimate.pdf}%}
% \subfigure[][]{\includegraphics{./Graph/KernelEstimateFFT.pdf}}\\
\caption{Cross-section of connectivity kernel estimate. The actual kernel is shown by the solid line, the mean kernel estimate (over 100 realisations) is shown by the dotted line and the $95~\%$ confidence interval is shown by the shaded grey region.}
\label{fig:KernelEstimates}
\end{figure}
The accuracy of the state estimates was evaluated by comparing the field reconstruction to the true field using the mean (over time) of the root mean square error (MRMSE) (over space). The RMSE of the field estimates over time for 100 realisations is shown in Figure~\ref{fig:RMSE}, with the $95~\%$ confidence interval. The error is consistently around 0.5 over time, which approximately $7.1$~\% of the voltages, which range over $(-3.29,3.77)$. The error in the reconstructed field is mostly due to the loss of high spatial frequency information due to the basis function decomposition; this is illustrated in Figure~\ref{fig:FieldEstimate}, which shows a snapshot at a single time instant of the true neural field, the estimated field and the corresponding error, which is the difference between the true and estimated neural fields. It can be seen in Figure~\ref{fig:FieldEstimate}(b) that the estimated field contains the basic structure of the true field (Figure~\ref{fig:FieldEstimate}(a)), but is missing the higher-frequency detail.
  \begin{figure}
   	\begin{center}
   		\includegraphics{./Graph/MRMSE.pdf} 
   	\end{center}
   	\caption{The mean of average RMSE of the estimated field over 100 realisations (solid line) and the 95~\% confidence interval (grey region).} 
\label{fig:RMSE}
   \end{figure}

@@

The closest competitor to the integro-difference equation based approach is perhaps the multivariate auto-regressive (MVAR) model. An advantage of the MVAR-based methods is that they allow for anisotropic connectivity, which is present in cortico-cortical fibres. A recent theoretical study~\cite{Jirsa2009} has demonstrated the importance of this multi-scale structure in generating the characteristic rhythms of the brain. A possible method of creating more precise models with this multi-scale structure maybe to estimate anisotropic cortico-cortical connectivity using a MVAR approach and intracortical connectivity using the methods proposed in this paper. In addition to estimating connectivity independently of the number of observations, the advantage of the methodology presented in this paper, is that the number of parameters required to model the system is considerably less, and the parameters have physical meaning.
